Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-08-20 paper1.tex
minor edits

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that
\[
%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
\]
Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.

@@

Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines are the variance predicted by Theorem~\ref{thm:normality}.  The dots and crosses show the results of Monte Carlo simulations.  Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.  This requires $G(0)$ to be known.  In practice $G(0)$ not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the estimator.  When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.

Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.  The figure depicts an interesting phenomenon.  When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.  When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator combining the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.  A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.  %When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.  
This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.  For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
\[
SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
\]
where the weight $\beta$ would be small when SNR is small and $\beta$ would be near $1$ when SNR is large.  For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.  One caveat is that the receiver would require knowledge about the noise distribution in order to correctly choose $\beta$.  We do not invesitigate this further here.

@@

Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.  Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).  This suggests the use of a weigthed objective function.  In order to select the weight it appears that the reciever would need some knowledge about the noise distribution.  It would be interesting to investigate methods for selecting, or estimating, the weight at the reciever.

>>>>

We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
\begin{align*}
\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
\end{align*}
Lemma~\ref{lem:empiricprocforrho} shows that 
\[
\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
\]
where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
\[
%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
\]
It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.  Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.  Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
\[
\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
\]
as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.

@@

Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines are the variance predicted by Theorem~\ref{thm:normality}.  The dots and crosses show the results of Monte-Carlo simulations.  Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.  This requires $G(0)$ to be known.  In practice $G(0)$ may not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.  When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.

Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.  The figure depicts an interesting phenomenon.  When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.  When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.  A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.  %When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.  
This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.  For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
\[
SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
\]
where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.  For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.  This would enable a rigorous theory for selection of $\beta$ at the reciever.  One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.  We do not invesitigate this further here.

@@

Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.  Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).  This suggests the use of a weigthed objective function, which would be an interesting topic for future research.
